---
title: "Gene-regulatory network of cardiomyocyte remodeling after myocardial infarction"
output: html_document
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this tutorial, we will use the integrated snRNA-seq and snATAC-seq data to infer a gene regulatory network to characterize cardiomyocyte remodeling. More details about
data integration can be found [here](https://costalab.github.io/scMEGA/articles/cardiomyocyte-preprocessing.html)

More specifically, we will:  

* Identify the trajectory for myofibroblast differentiation
* Select the relevant TFs and genes for this process
* Infer and visualize the gene regulatory network

Next, we load all necessary packages:
```{r}
suppressMessages(library(ArchR))
suppressMessages(library(Seurat))
suppressMessages(library(Signac))
suppressMessages(library(scMEGA))
suppressMessages(library(harmony))
suppressMessages(library(Nebulosa))
suppressMessages(library(BSgenome.Hsapiens.UCSC.hg38))
suppressMessages(library(JASPAR2020))
suppressMessages(library(TFBSTools))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(igraph))
suppressMessages(library(ggraph))
```

We first load the integrated data:
```{r, fig.height = 5, fig.width = 10, fig.align = "center"}
coembed.sub <- readRDS("./Cardiomyocyte/coembed.sub.rds")

cols <- ArchR::paletteDiscrete(coembed.sub@meta.data[, "RNA_snn_res.0.1"])

p <- DimPlot(coembed.sub, group.by = "RNA_snn_res.0.1", label = TRUE,
             reduction = "umap_harmony", shuffle = TRUE, split.by = "tech", 
             cols = cols) +
    xlab("UMAP1") + ylab("UMAP2")

p
```

## Trajectory analysis

### Cell pairing
We can match the cells between these two modalities. In other words, for each cell in, for example, snATAC-seq, we will find a cell from snRNA-seq data so that these two cells have the similar profiles. This is only necessary when each modality was performed independently. If snRNA-seq and snATAC-seq data was generated by multi-modal protocol, e.g., 10X multiome or SHARE-seq, this step can be skipped.

We here use the method proposed by [Kartha, Vinay K., et al.](https://www.biorxiv.org/content/10.1101/2021.07.28.453784v1.abstract) to match the cells.

```{r, fig.height = 5, fig.width = 10, fig.align = "center"}
df.pair <- PairCells(object = coembed.sub, reduction = "harmony",
                    pair.by = "tech", ident1 = "ATAC", ident2 = "RNA")
```

We can visualize the paired cells

```{r, fig.height = 5, fig.width = 10, fig.align = "center"}
sel_cells <- c(df.pair$ATAC, df.pair$RNA)
coembed.sub2 <- coembed.sub[, sel_cells]

options(repr.plot.height = 5, repr.plot.width = 10)
DimPlot(coembed.sub2, reduction = "umap_harmony", 
        group.by = "RNA_snn_res.0.1", split.by = "tech", cols = cols)
```

We next create a new Seurat object for there paired cells as if they are generated by single-cell
multimodal protocol.

```{r, fig.height = 5, fig.width = 10, fig.align = "center"}
obj.pair <- CreatePairedObject(df.pair = df.pair, 
                               object = coembed.sub2,
                               use.assay1 = "RNA", 
                               use.assay2 = "ATAC")

obj.pair
```

Finally, we infer a pseudo-time trajectory from SCARA5+ fibroblasts to myofibroblast using the approach from [ArchR](https://www.archrproject.com/). Here we modified the function to allow to take a Seurat object as input

```{r, fig.height = 5, fig.width = 10, fig.align = "center"}
obj.pair <- AddTrajectory(object = obj.pair, 
                          trajectory = c(2, 1, 0, 3),
                          group.by = "RNA_snn_res.0.1", 
                          reduction = "umap_harmony",
                          dims = 1:2, 
                          use.all = TRUE)
                          
# we only plot the cells that are in this trajectory
obj <- obj.pair[, !is.na(obj.pair$Trajectory)]

options(repr.plot.height = 5, repr.plot.width = 10)

p1 <- DimPlot(obj, reduction = "umap_harmony", 
              group.by = "RNA_snn_res.0.1", cols = cols)

p2 <- TrajectoryPlot(object = obj, 
                    reduction = "umap_harmony",
                    continuousSet = "blueYellow",
                    size = 1,
                   addArrow = FALSE) +
    xlab("UMAP 1") + ylab("UMAP 2") +
    ggtitle("Trajectory")

p1 + p2
```

## TF and gene selection
We next select candidate TFs and genes for building a meaningful gene regulatory network.

### Select TFs
To identify potential regulator (i.e., TFs), we first estimate an acitivty score for each TF in each cell. This is done by first performing motif matching and then computing deviation scores using [chromVAR](https://greenleaflab.github.io/chromVAR/index.html).

```{r}
# Get a list of motif position frequency matrices from the JASPAR database
pfm <- getMatrixSet(
  x = JASPAR2020,
  opts = list(collection = "CORE", tax_group = 'vertebrates', all_versions = FALSE)
)

# add motif information
obj <- AddMotifs(
  object = obj,
  genome = BSgenome.Hsapiens.UCSC.hg38,
  pfm = pfm,
    assay = "ATAC"
)

obj <- RunChromVAR(
  object = obj,
  genome = BSgenome.Hsapiens.UCSC.hg38,
    assay = "ATAC"
)

res <- SelectTFs(object = obj.trajectory, return.heatmap = TRUE,
cor.cutoff = 0.1)

df.cor <- res$tfs
ht <- res$heatmap
```

We can visualize the TF activity dynamic along the trajectory

```{r, fig.height = 6, fig.width = 10, fig.align = "center"}
draw(ht)
```

### Select genes

We will select relevant genes by first linking genes to peaks based on the corrleation between gene expression from the snRNA-seq data and peak accessibility from the snATAC-seq data along the inferred trajectory. This means that we only consider a gene to be a potential target if it can be assocaited to at least one peak.
```{r, fig.height = 6, fig.width = 10, fig.align = "center"}
res <- SelectGenes(object = obj,
                  labelTop1 = 0,
                  labelTop2 = 0)

df.p2g <- res$p2g
ht <- res$heatmap

draw(ht)
```

## Gene regulatory network inference and visualization

We here will try to predict a gene regulatory network based on the correlation of the TF binding activity as estimated from snATAC-seq and gene expression as measured by snRNA-seq along the trajectory.

```{r}
tf.gene.cor <- GetTFGeneCorrelation(object = obj, 
                                    tf.use = df.cor$tfs, 
                                    gene.use = unique(df.p2g$gene),
                                    tf.assay = "chromvar", 
                                    gene.assay = "RNA",
                                    trajectory.name = "Trajectory")
```

We can then visualize this correlation matrix by heatmap. Also, we can cluster the genes and TFs to identify different regulatory modules for the predefined sub-populations.

```{r, fig.height = 6, fig.width = 12, fig.align = "center"}
ht <- GRNHeatmap(tf.gene.cor, 
                 tf.timepoint = df.cor$time_point)
ht
```

To associate genes to TFs, we will use the peak-to-gene links and TF binding sites information. Specifically, if a gene is regulated by a peak and this peak is bound by a TF, then we consider this gene to be a target of this TF.

```{r, fig.height = 6, fig.width = 12, fig.align = "center"}
motif.matching <- obj@assays$ATAC@motifs@data
colnames(motif.matching) <- obj@assays$ATAC@motifs@motif.names
motif.matching <-
    motif.matching[unique(df.p2g$peak), unique(tf.gene.cor$tf)]


df.grn <- GetGRN(motif.matching = motif.matching, 
                 df.cor = tf.gene.cor, 
                 df.p2g = df.p2g)
```

Finally, we can visualize our network as the last step of this analysis

```{r viz_network, fig.height = 15, fig.width = 15, fig.align = "center"}
# define colors for nodes representing TFs (i.e., regulators)
df.cor <- df.cor[order(df.cor$time_point), ]
tfs.timepoint <- df.cor$time_point
names(tfs.timepoint) <- df.cor$tfs

# plot the graph, here we can highlight some genes
df.grn2 <- df.grn %>%
    subset(correlation > 0.5) %>%
    select(c(tf, gene, correlation)) %>%
    rename(weights = correlation)
    
p <- GRNPlot(df.grn2, 
             tfs.timepoint = tfs.timepoint,
             show.tf.labels = TRUE,
             seed = 42, 
             plot.importance = FALSE,
            min.importance = 2,
            remove.isolated = FALSE)

options(repr.plot.height = 20, repr.plot.width = 20)

print(p)
```

Save data
```{r, fig.height = 5, fig.width = 6, fig.align = "center"}
saveRDS(obj, "./Cardiomyocyte/coembed.trajectory.rds")
```

```{r}
# Check session information
sessionInfo()
```

